Supplementary Materials of Impact of IPTp Regimen on Pregnancy Outcomes

### 2025-05-03 products/manuscript/supplement/Supplementary-Material.qmd
update

```{r}
library(tidymodels)
library(readr)
library(here)

# 2. Read & split data
df <- read_csv(here("data","clean","PROMO_Data_clean.csv")) %>%
  mutate(adverse_birth_outcome = factor(adverse_birth_outcome, levels = c(0,1)))

set.seed(2025)
splits    <- initial_split(df, prop = 0.7, strata = adverse_birth_outcome)
train     <- training(splits)
test      <- testing(splits)
cv_folds  <- vfold_cv(train, v = 5, strata = adverse_birth_outcome)

# 3. Preprocessing recipe
ml_rec <- recipe(adverse_birth_outcome ~ 
                   age_at_enrollment_years +
                   gravidity +
                   total_malaria_episodes +
                   education_level +
                   study_arm +
                   normalized_malaria_rate,   # ← replaced here
                 data = train) %>%
  step_dummy(all_nominal_predictors()) %>%  # one‐hot encode
  step_zv(all_predictors()) %>%             # remove zero‐variance
  step_normalize(all_numeric_predictors())  # center & scale

# 4. Model specifications & workflows

## (a) LASSO‐penalized logistic regression
lasso_mod <- logistic_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet") %>%
  set_mode("classification")

lasso_wf  <- workflow() %>% 
  add_model(lasso_mod) %>% 
  add_recipe(ml_rec)

## (b) Random forest
rf_mod <- rand_forest(mtry = tune(), trees = 1000, min_n = tune()) %>%
  set_engine("ranger") %>%
  set_mode("classification")

rf_wf  <- workflow() %>% 
  add_model(rf_mod) %>% 
  add_recipe(ml_rec)

## (c) XGBoost
xgb_mod <- boost_tree(trees = tune(),
                      tree_depth = tune(),
                      learn_rate = tune()) %>%
  set_engine("xgboost") %>%
  set_mode("classification")

xgb_wf <- workflow() %>% 
  add_model(xgb_mod) %>% 
  add_recipe(ml_rec)

# 5. Tuning grids & 5‐fold CV
set.seed(2025)

lasso_grid <- grid_regular(penalty(), levels = 20)

rf_params <- parameters(rf_mod) %>%
  update(mtry = mtry(c(1, 6)))
rf_grid   <- grid_latin_hypercube(rf_params, size = 20)

xgb_params <- parameters(xgb_mod)
xgb_grid   <- grid_latin_hypercube(xgb_params, size = 30)

ml_metrics <- metric_set(roc_auc)

lasso_res <- tune_grid(lasso_wf, resamples = cv_folds,
                       grid = lasso_grid, metrics = ml_metrics)

rf_res    <- tune_grid(rf_wf,    resamples = cv_folds,
                       grid = rf_grid,    metrics = ml_metrics)

xgb_res   <- tune_grid(xgb_wf,   resamples = cv_folds,
                       grid = xgb_grid,   metrics = ml_metrics)

library(dplyr)
library(knitr)

# 6) Pull the metrics into plain tibbles
lasso_metrics <- collect_metrics(lasso_res)
rf_metrics    <- collect_metrics(rf_res)
xgb_metrics   <- collect_metrics(xgb_res)

# 7) Summarize AUC for each model
lasso_auc <- lasso_metrics %>%
  filter(.metric == "roc_auc") %>%
  dplyr::summarise(
    Mean = mean(mean),
    SE   = mean(std_err)
  ) %>%
  mutate(Model = "LASSO")

rf_auc <- rf_metrics %>%
  filter(.metric == "roc_auc") %>%
  dplyr::summarise(
    Mean = mean(mean),
    SE   = mean(std_err)
  ) %>%
  mutate(Model = "Random Forest")

xgb_auc <- xgb_metrics %>%
  filter(.metric == "roc_auc") %>%
  dplyr::summarise(
    Mean = mean(mean),
    SE   = mean(std_err)
  ) %>%
  mutate(Model = "XGBoost")

# 8) Combine into one data frame and print as a plain table
table6 <- bind_rows(lasso_auc, rf_auc, xgb_auc) %>%
  dplyr::select(Model, Mean, SE)

print(table6)

>>>>

```{r}
library(dplyr)
library(knitr)

# 6) Collect tuning metrics
lasso_metrics <- collect_metrics(lasso_res)
rf_metrics    <- collect_metrics(rf_res)
xgb_metrics   <- collect_metrics(xgb_res)

# 7) Summarize AUC for each model
lasso_auc <- lasso_metrics %>%
  filter(.metric == "roc_auc") %>%
  summarise(
    Mean = mean(mean),
    SE   = mean(std_err)
  ) %>%
  mutate(Model = "LASSO")

rf_auc <- rf_metrics %>%
  filter(.metric == "roc_auc") %>%
  summarise(
    Mean = mean(mean),
    SE   = mean(std_err)
  ) %>%
  mutate(Model = "Random Forest")

xgb_auc <- xgb_metrics %>%
  filter(.metric == "roc_auc") %>%
  summarise(
    Mean = mean(mean),
    SE   = mean(std_err)
  ) %>%
  mutate(Model = "XGBoost")

# 8) Combine into one data frame
table6 <- bind_rows(lasso_auc, rf_auc, xgb_auc)

# 9) Render with kable for Word, explicitly namespacing select
kable(
  table6 %>% dplyr::select(Model, Mean, SE),
  format  = "pandoc",
  digits  = 3,
  caption = "Table 6: Cross‑validated AUC (mean ± SE) for Each Model"
)
